### 2019-07-04 poster/poster.tex
Add slides with makefile

% Gemini theme
% https://github.com/anishathalye/gemini

\documentclass[final]{beamer}

% ====================
% Packages
% ====================

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[size=custom,width=120,height=72,scale=1.0]{beamerposter}
\usetheme{gemini}
\usecolortheme{gemini}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tikz}
 \usepackage{ragged2e}
\usepackage{pgfplots}
\addtobeamertemplate{block begin}{}{\justifying}  %new code

% ====================
% Lengths
% ====================

% If you have N columns, choose \sepwidth and \colwidth such that
% (N+1)*\sepwidth + N*\colwidth = \paperwidth
\newlength{\sepwidth}
\newlength{\colwidth}
\setlength{\sepwidth}{0.025\paperwidth}
\setlength{\colwidth}{0.3\paperwidth}

\newcommand{\separatorcolumn}{\begin{column}{\sepwidth}\end{column}}

% ====================
% Title
% ====================
\title{Software Clusterings with Vector Semantics and the Call Graph}
\author[Marios Papachristou]{Marios Papachristou}

\institute[BaLab]{BALab, Athens University of Economics and Bussiness \\ National Technical University of Athens \\ Advisor: Prof. Diomidis Spinellis}
\date[ESEC/FSE 2019 SRC]{Student Research Competition \\ ESEC/FSE August 2019 \\ Estonia}

% ====================
% Body
% ====================

\begin{document}

\begin{frame}[t]
\begin{columns}[t]
\separatorcolumn

\begin{column}{\colwidth}

  \begin{alertblock}{ABSTRACT}

  	\textbf{Purpose --}  Propose a method to determine a software’s modules
without knowledge of its structure, and empirically validate the
method’s performance.

	\textbf{Method --} Cluster files by combining \textbf{document embeddings}, generated with the \textbf{Doc2Vec} algorithm, and the \textbf{call graphs}, provided
by Static Graph Analyzers to an \textbf{augmented graph}. Use of the Louvain Algorithm to iteratively exhibit its community structure and propose a module–level clustering.

	\textbf{Results --} Our method performs better in terms of stability, 			authoritativeness and extremity over other state-of-the-art clustering
methods proposed in literature and is able to decently recover the
ground truth clustering of the Linux Kernel.

	\textbf{Conclusions --} Semantic information from vector semantics and
the call graph can produce accurate results for software clusterings
of large systems.
  \end{alertblock}

  \begin{block}{Problem Statement}

	\begin{itemize}
    \item The architecture of a software system is the most fundamental realization of it.
    \item When there is no specific definition of it, we can attempt to recover it.
    \item One particular problem is the \textbf{clustering of its components into modules}. 
    \item Many methods exist in literature such as agglomerative clustering baselines as well as more sophisticated algorithms, like LIMBO and ACDC. 
\end{itemize}

  \end{block}

  \begin{alertblock}{Purpose}

\begin{enumerate}
    \item \textbf{Provide} a method for software clusterings through \textbf{vector semantics} and the \textbf{call graph} to determine software's modules without knowledge of its structure.
    \item Empirically validate our method performance on the \textbf{Linux Kernel Codebase}.

    \item Compare it against state-of-the-art methods (ACDC \cite{acdc}, LIMBO \cite{LIMBO}) and agglomerative clustering methods (agglomerative clustering \cite{average, complete, ward}). 
\end{enumerate}

  \end{alertblock}

\begin{block}{Related Work}

\begin{itemize}
\justifying

\item Onaiza Maqbool and Haroon A. Babri \cite{maqbool_overview} present an
overview of the various approaches taken towards hierarchical clustering algorithms
for software systems. In their study, they compare many clustering algorithms
such as \textbf{ACDC}, \textbf{LIMBO} and other Traditional Clustering Algorithms such as Single
Linkage, Complete Linkage and Weighted Linkage presenting their efficiency regarding 
multiple feature sets.

\item A \textbf{semantics-based architectural view} of the system, as discussed in reference
\cite{large_study} reveals significant aspects of a software system and its change over 
time, which suggests that semantic-based approaches should be followed for gaining better understandings of a software system.

\item The \textbf{ACDC} algorithm leans toward to software components comprehension 
based on subsystem patterns. Their approach considers an initial structure of the system, 
without taking into account semantics, and tries to build comprehensive clusterings of the given ground structure.

\item Andritsos and Tzerpos in reference \cite{LIMBO} present an Information-Theoretic 
approach of Software Clustering Algorithms by developing \textbf{LIMBO} which clusters modules upon inserting their Distributional Cluster Features 
to a B+-tree variant and then applying the \textbf{Agglomerative Information Bottleneck} algorithm.
\end{itemize}

\end{block}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

\begin{alertblock}{Our Method}

\heading{Preprocessing of the codebase \& Training a Doc2Vec Model}
First of all, we tokenize the source code and remove the stop-words. 
After that, for each token, we split it into its constituent parts using dynamic programming \cite{wordninja} and lemmatize 
each individual sub-token via using the \textbf{NLP} package spaCy \cite{spacy}. 
For example, the method \texttt{\_\_zone\_seqlock\_init} corresponds to \texttt{zone}, \texttt{seqlock} and \texttt{init}, 
\texttt{inprogress} is split into \texttt{in} and \texttt{progress} and \texttt{literals} becomes \texttt{literal}.
We, then, train a Doc2Vec model using Gensim \cite{gensim}.

\heading{Extracting the Call Graph(s)}

\begin{enumerate}
\item The codebase is also processed by a Static Graph Analyzer. 
In our approach, focusing on C projects, we use \textbf{CScout} \cite{cscout} in order to generate the directed call graph through function calls between the files. 
Each source code file is assigned to a module. 
Modules can be user-defined or automatically generated by their respective directories at a desired directory tree depth.

\item The edges have the normalized cosine similarities as weights $w(i,j) = (1 + \cos(\mathbf x_i, \mathbf x_j)) / 2$.
\end{enumerate}

\heading{Obtaining a clustering}
Then we run the Louvain Clustering Algorithm \cite{louvain} in
order to obtain the software clustering via maximizing the modularity function with a greedy approach. 
The Louvain Method is a greedy optimization method that (approximately) maximizes the modularity function of a graph $H(V, E)$

$$Q(H) = \frac {1}{2m} \sum_{(i, j) \in E(H)} \left ( w(i, j) - \frac {k(i) k(j)} {2m} \right )$$

where $m = \sum_{(i, j) \in E} w(i,j)$ and $k(i) = \sum_{j \in \mathrm{in}(i)} w(i, j)$ . In case we want to consider edge directionality, we do a bipartite transformation, detect communities and
merge the results with a union-find data structure as stated in reference \cite{malliaros}.

\heading{Decision Rationale for Evaluation System}

We have chosen to use Linux as a codebase to evaluate our method on since

\begin{enumerate}

\item it is a large and complex system with 27 years of continuous development. 
\item it spans $\sim$20.3 million lines of source code at the time of writing.
\item  it is easy to establish a ground truth due to its clear structure and construct test-cases for evaluations.

\item it is used in related studies \cite{acdc, evaluation} as a reference system for evaluation.

\end{enumerate}

\heading{Ground Truth}
As ground truth, we have used the first level directories as a target clustering and as input, we have considered 
the modules of the one-top directories. 
For example, the source code file \texttt{drivers/net/ieee802154\-/mcr20a.c} has a ground truth value of \texttt{drivers} 
and it is considered under the same module as every \texttt{.c} and \texttt{.h} file under \texttt{drivers/net/ieee802154}.

\heading{Experimental Evaluation}

\begin{itemize}

\item Compare our method against agglomerative clustering baselines (average, complete, and Ward linkage) and state-of-the-art methods (LIMBO and ACDC).
\item Use the MoJo distance metric to calculate the distance between the proposed clustering and ground truth.
\item Evaluate clusterings \textbf{stability}, \textbf{authoritativeness} and \textbf{extremity}.

\end{itemize}

\end{alertblock}

\end{column}

\separatorcolumn

\begin{column}{\colwidth}

  \begin{block}{Results}

    \begin{table}
      \centering
      \small
\begin{tabular}{lrrrrrrr}
    \toprule
    \textbf{Algorithm} & \textbf{Feat. Dim}  & $n_c$ & \textbf{Range} & $\bar x$ & $\sigma$ & \textbf{Median} & \textbf{MoJo D.} \\
    \midrule
    \textbf{ACDC}  & -- & 9055 & 1 -- 4245 & 5 & 46 & 2 & 33694\\
    Average Linkage  & \emph{300} & \emph{21} & 1--3406 & 163 & 725 & 1 & 2092 \\
    Complete Linkage  & \emph{300} & \emph{21} & 1--1529 & 163 & 412 & 19 & 1710 \\
    \textbf{LIMBO}  ($B=100, \; S = \infty$) & 12317 &\emph{21} & 50--1810 & 163 & 375 & 50  & 1482 \\

    Ward Linkage & \emph{300} & \emph{21} & 21--948 & 163 & 223 & 70 & 1138 \\

    \textbf{SADE} & \emph{300} & 10 ($\pm$ 2)  & 2 ($\pm$ 0) -132 ($\pm$ 13) & 64 ($\pm$ 4) & 40 ($\pm$ 4) & 65 ($\pm$ 10) & 243 ($\pm$ 1)  \\
    \textbf{SADE} (Directed) & \emph{300} & 5 ($\pm$ 2) & 1 ($\pm$ 1) - 614 ($\pm$ 1) & 141 ($\pm$ 39) & 253 ($\pm$ 25) & 2 ($\pm$ 0.3)  & 237 ($\pm$ 2) \\
    \midrule
    Ground Truth & -- & 21 & 1--1348 & 163 & 341 & 11.0 & -- \\
    \bottomrule
  \end{tabular}

      \caption{Experimental Results for Linux Kernel 4.21 Codebase. User-defined parameters are in italics.}
    \end{table}

	\begin{itemize}
		\justifying
    \item \textbf{Authoritativeness -- } \textbf{Surpass} all clustering algorithms in terms of MoJo Distance Metric.
    \item \textbf{Extremity -- } Production of \textbf{balanced clusterings}.
    \item \textbf{Stability -- } Production of \textbf{stable clusterings}.
    \item Results were produced \textbf{without knowing the number of clusters of the ground truth a priori}.
    \item Provide a \textbf{simplistic approach} to software clustering combining vector semantics and the call graph.
	\end{itemize}

  \end{block}

	\begin{alertblock}{Conclusions}
	\begin{enumerate}
	\justifying
    \item Use of \textbf{vector semantics} and the \textbf{call graph} to produce meaningful clusterings.
    \item Performing our study on a very large system (Linux) gives us \textbf{further insight} on the nature of software itself.
    \item \textbf{Outperform} state-of-the-art and baseline methods in terms of authoritativeness and extremity.
    \item Produce \textbf{stable} and \textbf{balanced} clusterings. 
\end{enumerate}

	\end{alertblock}

  \begin{block}{References}

    \scriptsize{\bibliographystyle{plain}\bibliography{references}}

  \end{block}

\end{column}

\separatorcolumn
\end{columns}
\end{frame}

\end{document}