### 2019-11-29 extended-manuscript/extended-manuscript.tex
Setup outline for extended paper

\documentclass[a4paper, 11pt]{article}
\usepackage[margin=0.7in]{geometry}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage[english]{babel}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}

\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}

\usepackage{graphicx}
\renewcommand\vec{\mathbf}

\title{\textbf{Software Clusterings with Vector Semantics and the Call Graph}\footnote{The research described has been carried out as part of the CROSSMINER Project, which has received funding from the European Union's Horizon 2020 Research and Innovation Programme under grant agreement No. 732223}\;\footnote{This work serves as an extended version of \cite{sade}}}

\author{
  Marios Papachristou\footnote{Also affiliated with the National Technical University of Athens} \\
  Business Analytics Lab, AUEB \\
  \texttt{papachristoumarios@gmail.com}
  \and
  Diomidis Spinellis \\
  Business Analytics Lab, AUEB \\
  \texttt{dds@aueb.gr}
}

\date{}

\begin{document}

\maketitle

\section{Introduction}

\section{Method}

\section{Evaluation}

\section{Results and Discussion}

\section{Related Work}

\section{Conclusion}

\newpage

\nocite{*}
\bibliographystyle{alpha}
\bibliography{extended-references}

\end{document}